A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution \thanks{A preliminary version of this work has appeared in the Proceedings of the 26th International Symposium on Graph Drawing and Network Visualization (GD 2018). This work is supported by the DFG grant Ka812/17-1}

### 2018-10-04 paper/journal.tex
some more work on the journal
Table 2 -> update with Henry
current version

\usepackage{amsmath,amssymb}
\usepackage{graphicx,pgfplots}
\usepackage[font=small]{subfig,caption}
%\usepackage{algorithm2e}
\usepackage{todonotes,lineno,paralist,soul}
\usepackage{rotating,hyperref,multicol}
\usepackage{booktabs,tabularx}
\graphicspath{{figures/}}

@@

The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$. It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced by Section~\ref{sec:relatedwork}.

@@

\paragraph{Structure of the paper:}
The remainder of this paper is structured as follows. Section~\ref{sec:relatedwork} overviews related works. Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project. We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.

% ==================================================================
\section{Related Work}
\label{sec:relatedwork}
% ==================================================================

As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem. Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}. The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge. Actually, several relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}. Deciding, however, whether a $1$-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}. Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.

@@

@@

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2017.}
\centering
\begin{tabular}{c|c|c|c|c}
\toprule
~~Graph~~ & ~~CoffeeVM~~ & ~~TuebingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
\midrule
1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$ & 89.99$^\circ$\\
2 & 88.23$^\circ$  & 42$^\circ$ & 88.21$^\circ$ & 88.70$^\circ$\\
3 & 90.00$^\circ$  & 89$^\circ$ & 87.86$^\circ$ & 89.95$^\circ$\\
4 & 88.97$^\circ$  & 89$^\circ$ & 77.13$^\circ$ & 89.05$^\circ$\\
5 & 80.40$^\circ$  & 30$^\circ$ & 78.68$^\circ$ & 86.96$^\circ$\\
6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$ & 89.96$^\circ$\\
7 & 56.53$^\circ$  & 34$^\circ$ & 55.77$^\circ$ & 63.62$^\circ$\\
8 & 84.95$^\circ$  & 61$^\circ$ & 81.18$^\circ$ & 89.28$^\circ$\\
9 & 59.88$^\circ$  & 9$^\circ$  & 54.63$^\circ$ & 88.20$^\circ$\\
10 & 20.97$^\circ$ & 4$^\circ$  & 23.60$^\circ$ & 23.72$^\circ$\\
11 & 46.68$^\circ$ & 6$^\circ$  & 57.00$^\circ$ & 72.00$^\circ$\\
12 & 36.47$^\circ$ & 5$^\circ$  & 26.24$^\circ$ & 35.86$^\circ$\\
13 & 25.45$^\circ$ & 4$^\circ$  & 22.43$^\circ$ & 33.68$^\circ$\\
14 & 33.52$^\circ$ & 5$^\circ$  & 29.69$^\circ$ & 43.08$^\circ$\\
15 & 20.51$^\circ$ & 4$^\circ$  & 13.51$^\circ$ & 29.18$^\circ$\\
16 & 19.83$^\circ$ & 3$^\circ$  & 12.13$^\circ$ & 26.12$^\circ$\\
\bottomrule
\end{tabular}
\label{tab:gdContest2017}
\end{table*}

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2018.}
\centering
\begin{tabular}{c|c|c|c}
\toprule
~~Graph~~ & ~~TuebingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
\midrule
1 & 90.00$\circ$   & 77$^\circ$ & 89.99$^\circ$\\
2 & 88.23$^\circ$  & 42$^\circ$ & 88.70$^\circ$\\
3 & 90.00$^\circ$  & 89$^\circ$ & 89.95$^\circ$\\
4 & 88.97$^\circ$  & 89$^\circ$ & 89.05$^\circ$\\
5 & 80.40$^\circ$  & 30$^\circ$ & 86.96$^\circ$\\
6 & 90.00$^\circ$  & 78$^\circ$ & 89.96$^\circ$\\
7 & 56.53$^\circ$  & 34$^\circ$ & 63.62$^\circ$\\
8 & 84.95$^\circ$  & 61$^\circ$ & 89.28$^\circ$\\
9 & 59.88$^\circ$  & 9$^\circ$  & 88.20$^\circ$\\
10 & 20.97$^\circ$ & 4$^\circ$  & 23.72$^\circ$\\
11 & 46.68$^\circ$ & 6$^\circ$  & 72.00$^\circ$\\
12 & 36.47$^\circ$ & 5$^\circ$  & 35.86$^\circ$\\
13 & 25.45$^\circ$ & 4$^\circ$  & 33.68$^\circ$\\
14 & 33.52$^\circ$ & 5$^\circ$  & 43.08$^\circ$\\
\bottomrule
\end{tabular}
\label{tab:gdContest2018}
\end{table*}

@@

% ==================================================================
\subsection{Graph Drawing Contests 2018 and 2019}
\label{ssec:contest}
% ==================================================================

\myparagraph{Graph Drawing Contest 2017} In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
%
Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.

\myparagraph{Graph Drawing Contest 2018} In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper). The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''. Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.

It is eye-catching that our team was the winning one; team \texttt{CoffeeVM+} took the second position of the contest (out of a total of four teams that participated to the contest). It is worth noting that for almost all $14$ challenge graphs our team achieved significantly better results than those of the second team \texttt{CoffeeVM+}, which clearly indicates that our new approach is very competitive (and definitely better than our previous,  as also evidenced by the experimental evaluation).

>>>>

\usepackage{amsmath,amssymb}
\usepackage{graphicx,color,pgfplots}
\usepackage[font=small]{subfig,caption}
\usepackage{todonotes,lineno,paralist,soul}
\usepackage{rotating,hyperref,multicol}
\usepackage{booktabs,tabularx,colortbl}
\graphicspath{{figures/}}

@@

The latter result is an indication that the problem of finding drawings with high crossing resolution might also be difficult, even though, formally, its complexity has not been settled yet for values of the crossing resolution smaller than $90^\circ$. It is also worth noting that the literature is significantly more limited, when restricting the crossing resolution to be smaller than $90^\circ$, as also evidenced from our related work section.

@@

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2017.}
\centering
\begin{tabular}{c|c|c|c|c}
\toprule
\rowcolor{gray!30!}~~Graph~~ & ~~CoffeeVM~~ & ~~T\"ubingenMidnight~~ & ~~Time restricted~~ & ~~Our best~~ \\
\midrule
\rowcolor{gray!30!} 1  & 90$^\circ$  & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 2  & 88$^\circ$  & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 3  & 90$^\circ$  & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 4  & 89$^\circ$  & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 5  & 80$^\circ$  & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
\rowcolor{gray!00!} 6  & 90$^\circ$  & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!30!} 7  & 57$^\circ$  & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
\rowcolor{gray!00!} 8  & 85$^\circ$  & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 9  & 60$^\circ$  & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
\rowcolor{gray!00!} 10 & 21$^\circ$  & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
\rowcolor{gray!30!} 11 & 47$^\circ$  & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
\rowcolor{gray!00!} 12 & 36$^\circ$  & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
\rowcolor{gray!30!} 13 & 25$^\circ$  & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
\rowcolor{gray!00!} 14 & 34$^\circ$  & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
\rowcolor{gray!30!} 15 & 21$^\circ$  & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
\rowcolor{gray!00!} 16 & 20$^\circ$  & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
\bottomrule
\end{tabular}
\label{tab:gdContest2017}
\end{table*}

% ==================================================================
\paragraph{Related Work:}
%\label{sec:relatedwork}
% ==================================================================
As already mentioned, the study of the crossing resolution maximization problem has mainly focused on (properties of) RAC graphs, that is, on the optimal case of the crossing resolution maximization problem. Their study was initiated by Didimo et al.~\cite{DBLP:journals/tcs/DidimoEL11}, who showed that an $n$-vertex RAC graph has at most $4n-10$ edges~\cite{DBLP:journals/tcs/DidimoEL11}, while deciding whether a graph is RAC is NP-hard~\cite{DBLP:journals/jgaa/ArgyriouBS12}. The maximally-dense RAC graphs are 1-planar~\cite{DBLP:journals/dam/EadesL13}, i.e., they can be drawn with at most one crossing per edge. Actually, several relationships between the class of RAC graphs and subclasses of 1-planar graphs are known~\cite{DBLP:journals/dam/BachmaierBHNR17,DBLP:journals/tcs/BrandenburgDEKL16}. Deciding, however, whether a $1$-planar graph is RAC is NP-hard~\cite{DBLP:journals/tcs/BekosDLMM17}. Note that the problem of finding RAC drawings has also been studied in the presence of bends~\cite{DBLP:journals/jgaa/AngeliniCDFBKS11,DBLP:journals/comgeo/ArikushiFKMT12,DBLP:journals/tcs/DidimoEL11,DBLP:journals/mst/GiacomoDLM11} and by imposing restrictions on the degree~\cite{DBLP:conf/s-egc/AngeliniBDFHKLL11}, the structure~\cite{DBLP:journals/ipl/DidimoEL10} and the drawing~\cite{DBLP:journals/algorithmica/GiacomoDEL14,DBLP:conf/wg/HongN15} of the graph.

@@

% ==================================================================
\paragraph{Motivation:}
%\label{sec:motivation}
% ==================================================================

%\myparagraph{Graph Drawing Contest 2017} 
In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al. and by Huang et al.). 
%
Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2018.}
\centering
\begin{tabular}{c|c|c|c|c}
\toprule
\rowcolor{gray!30!} ~~Graph~~ & ~~Grid size~~ & ~~T\"ubingenColdShower~~ & ~~CoffeeVM+~~ & ~~Our best~~ \\
\midrule
\rowcolor{gray!30!} 1  & 100 $\times$ 100             & 87$^\circ$ & 58$^\circ$ & 89.99$^\circ$\\
\rowcolor{gray!00!} 2  & 100 $\times$ 100             & 90$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!30!} 3  & 150 $\times$ 150             & 88$^\circ$ & 85$^\circ$ & 89.95$^\circ$\\
\rowcolor{gray!00!} 4  & 200 $\times$ 200             & 88$^\circ$ & 60$^\circ$ & 89.05$^\circ$\\
\rowcolor{gray!30!} 5  & 200 $\times$ 200             & 71$^\circ$ & 48$^\circ$ & 78$^\circ$\\
\rowcolor{gray!00!} 6  & 200 $\times$ 200             & 89$^\circ$ & 71$^\circ$ & 89.96$^\circ$\\
\rowcolor{gray!30!} 7  & 250 $\times$ 250             & 80$^\circ$ & 43$^\circ$ & 63.62$^\circ$\\
\rowcolor{gray!00!} 8  & 10,000 $\times$ 10,000       & 34$^\circ$ & 27$^\circ$ & 43$^\circ$\\
\rowcolor{gray!30!} 9  & 10,000 $\times$ 10,000       & 77$^\circ$ & 60$^\circ$ & 83$^\circ$\\
\rowcolor{gray!00!} 10 & 30,000 $\times$ 30,000       & 13$^\circ$ & 19$^\circ$ & 24$^\circ$\\
\rowcolor{gray!30!} 11 & 50,000 $\times$ 50,000       & 11$^\circ$ & 22$^\circ$ & 72.00$^\circ$\\
\rowcolor{gray!00!} 12 & 100,000 $\times$ 100,000     & 74$^\circ$ &  9$^\circ$ & 80$^\circ$\\
\rowcolor{gray!30!} 13 & 1,000,000 $\times$ 1,000,000 &  2$^\circ$ &  2$^\circ$ & 33.68$^\circ$\\
\rowcolor{gray!00!} 14 & 1,000,000 $\times$ 1,000,000 &  3$^\circ$ & 11$^\circ$ & 11$^\circ$\\
\bottomrule
\end{tabular}
\label{tab:gdContest2018}
\end{table*}

%\myparagraph{Graph Drawing Contest 2018} 
In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper). The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''. Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.

Note that in year 2018, the organizers of the contest posed one additional constraint, namely, the drawing area was restricted to a certain size, which was different for each graph; see column ``Grid size'' in Table~\ref{tab:gdContest2018}. It is eye-catching that our team was the winning one; team \texttt{CoffeeVM+} took the second position of the contest (out of a total of four teams that participated to the contest). It is worth noting that for almost all $14$ challenge graphs our team achieved significantly better results than those of the second team \texttt{CoffeeVM+}, which clearly indicates that our new approach is very competitive (and definitely better than our previous,  as also evidenced by the experimental evaluation).

\paragraph{Structure of the paper:}
The remainder of this paper is structured as follows. 
%Section~\ref{sec:relatedwork} overviews related works. 
Our algorithm is presented in detail in Section~\ref{sec:algorithm} and is experimentally evaluated against the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} in Section~\ref{sec:experiments}, where we also discuss our insights from this project. We conclude in Section~\ref{sec:conclusions} with a discussion of useful observations and insights that we obtained while working on this project.

@@

@@
